
# Models with Random Effects

## Import Data

```{r comment = ""}

#  "\t"  INDICATES SEPARATED BY TABLES  ;
taxprep  = read.table("TXTData/TaxPrep.txt", sep ="\t", quote = "",header=TRUE)

# taxprep=read.table(choose.files(), header=TRUE, sep="\t")
```

 Data for this study are from the Statistics of Income (SOI) Panel of Individual Returns, a part of the Ernst and Young/University of Michigan Tax Research Database. The SOI Panel represents a simple random sample of unaudited individual income tax returns filed for tax years 1979-1990. The data are compiled from a stratified probability sample of unaudited individual income tax returns, Forms 1040, 1040A and 1040EZ, filed by U.S. taxpayers. The estimates that are obtained from these data are intended to represent all returns filed for the income tax years under review. All returns processed are subjected to sampling except tentative and amended returns.
 
 Variable| Description
---- | --------------------------------------------------------- 
MS |is an indicator variable of the taxpayer's marital status. It is coded one if the taxpayer is married and zero otherwise.
HH | is an indicator variable, one if the taxpayer is a head of household and zero otherwise.
DEPEND | is the number of dependents claimed by the taxpayer.
AGE | is the presence of an indicator for age 65 or over.
F1040A | is an indicator variable of the taxpayer’s filing type. It is coded one if the taxpayer uses Form 1040A and zero otherwise.
F1040EZ | is an indicator variable of the taxpayer’s filing type. It is coded one if the taxpayer uses Form 1040EZ and zero otherwise.
TPI | is the sum of all positive income line items on the return. is a marginal tax rate.
TXRT | is a marginal tax rate。 It is computed on TPI less exemptions and the standard deduction.
MR | is an exogenous marginal tax rate. It is computed on TPI less exemptions and the standard deduction.
EMP | is an indicator variable, one if Schedule C or F is present and zero otherwise. Self-employed taxpayers have greater need for professional assistance to reduce the reporting risks of doing business.
PREP | is a variable indicating the presence of a paid preparer.
TAX | is the tax liability on the return.
SUBJECT |  Subject identifier, 1- 258. 
TIME | Time identifier, 1-5.
LNTAX | is the natural logarithm of the tax liability on the return.
LNTPI | is the natural logarithm of the sum of all positive income line items on the return.

## Example 3.2: Income Tax Payments (Page 81)

 In this section, we study the effects that an individual’s economic and demographic characteristics have on the amount of income tax paid. Specifically, the response of interest is `LNTAX`, defined as the natural logarithm of the liability on the tax return. 

### Table 3.2. Averages of binary variables

The binary variables in Table 3.2 indicate that over half the sample is married (MS) and approximately half the sample uses a paid preparer (PREP).

```{r comment = ""}
library(nlme)
gsummary(taxprep[, c("MS", "HH", "AGE", "EMP", "PREP")], groups=taxprep$TIME, FUN=mean)
```


### TABLE 3.3 - Summary statistics for continuous variables

Tables 3.2 and 3.3 describe the basic taxpayer characteristics used in our analysis. The summary statistics for the other nonbinary variables are in Table 3.3.

```{r comment = ""}

summary(taxprep[, c("DEPEND", "LNTPI", "MR", "LNTAX")]) #summary does not provid standard deviation
```

 Standard deviation of some variables.

```{r comment = ""}
#Standard Deviation
var<-var(taxprep[, c("DEPEND", "LNTPI", "MR", "LNTAX")])
sqrt(diag(var))
```

### TABLE 3.4 - Averages by level of binary explanatory variable
 
 To explore the relationship between each indicator variable and logarithmic tax, Table 3.4 presents the average logarithmic tax liability by level of indicator variable. This table shows that married filers pay greater tax, head-of-household filers pay less tax, taxpayers 65 or over pay less, taxpayers with self-employed income pay less, and taxpayers who use a professional tax preparer pay more.

```{r comment = ""}
library(Hmisc)
summarize(taxprep$LNTAX, taxprep$MS, mean) 
summarize(taxprep$LNTAX, taxprep$HH, mean)
summarize(taxprep$LNTAX, taxprep$AGE, mean)
summarize(taxprep$LNTAX, taxprep$EMP, mean)
summarize(taxprep$LNTAX, taxprep$PREP, mean)

# TABLE counts of BINARY EXPLANATORY VARIABLE
# CREATE CATEGORICAL VARIABLE
taxprep$MSF=taxprep$MS
taxprep$HHF=taxprep$HH
taxprep$AGEF=taxprep$AGE
taxprep$EMPF=taxprep$EMP
taxprep$PREPF=taxprep$PREP
table(taxprep$MSF)
table(taxprep$HHF)
table(taxprep$AGEF)
table(taxprep$EMPF)
table(taxprep$PREPF)
```

### TABLE 3.5 - Correlation for continous variables
 Table 3.5 summarizes basic relations among logarithmic tax and the other nonbinary explanatory variables. Both `LNTPI` and `MR` are strongly correlated with logarithmic tax whereas the relationship between `DEPEND` and logarithmic tax is positive, yet weaker. Table 3.5 also shows that `LNTPI` and `MR` are strongly positively correlated.
 
```{r comment = ""}
cor(taxprep[,c("LNTAX", "DEPEND", "LNTPI", "MR")])
```
 

### FIGURE 3.2: Basic added variable plot (y vs. x)

Moreover, both the mean and median marginal tax rates (`MR`) are decreasing, although mean and median tax liabilities (`LNTAX`) are stable (see Figure 3.2). These results are consistent with congressional efforts to reduce rates and expand the tax base through broadening the definition of income and eliminating deductions.

```{r comment = ""}
#CREATE CATEGORICAL VARIABLE
taxprep$SUBJECT1=factor(taxprep$SUBJECT)
lntax.lm = lm(LNTAX ~ SUBJECT1, data=taxprep)
lntpi.lm = lm(LNTPI ~ SUBJECT1, data=taxprep)
taxprep$Resid1=residuals(lntax.lm)
taxprep$Resid2=residuals(lntpi.lm)
plot(Resid1 ~ Resid2, data=taxprep, xaxt="n", yaxt="n", ylab="", xlab="")
axis(2, at=seq(-8, 7, by=2), las=1, font=10, cex=0.005, tck=0.01)
axis(2, at=seq(-8, 8, by=0.2), lab=F, tck=0.005)
axis(1, at=seq(-8,4, by=2), font=10, cex=0.005, tck=0.01)
axis(1, at=seq(-8, 4, by=0.2), lab=F, tck=0.005)
mtext("Residuals from LNTAX", side=2, line=-7, at=7.5, font=10, cex=1, las=1)
mtext("Residuals from LNTPI", side=1, line=3, at=-2, font=10, cex=1)
```

### DISPLAY 3.1 - Error components model
 
 The estimated model appears in Display 3.1, from a fit using the statistical package SAS. Display 3.1 shows that HH, EMP, LNTPI, and MR are statistically significant variables that affect LNTAX. Somewhat surprisingly, the PREP variable was not statistically significant.
 
```{r comment = ""}
random<-lme(LNTAX~MS+HH+AGE+EMP+PREP+LNTPI+DEPEND+MR, data=taxprep, random=~1|SUBJECT, method="ML")
## NOTE* THE DEFAULT METHOD IN lme IS "REML"
summary(random)
```
 
## SECTION 3.3 - Random coefficients model

```{r comment = ""}

#randomcoeff<-lme(LNTAX~MS+HH+AGE+EMP+PREP+LNTPI+DEPEND+MR, data=taxprep, random=~1+MS+HH+AGE+EMP+PREP+LNTPI+DEPEND+MR|SUBJECT, method="ML") 
# NOTE*:It takes forever to run the estimation, in the end a warning messaged was given. 
# No estimation result was produced. 
# The reason is due to the fact that in SAS, the method of mivque0 allows estimation for this model, in R this method is not readily available to be coded.

```
